---
title: "computingchallengeI"
author: "brockad"
date: "2020-09-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
This is the basic computing I challenge from [qBio6]()

```{r Package loading}
library(tidyverse)
```

```{r File download and data tidying}
temp <- tempfile()
download.file("https://raw.githubusercontent.com/jnovembre/BSD-QBio6/master/tutorials/basic_computing_1/nobel_nominations.csv",temp, mode = "wb")
  
nobel_nominations <- read_csv(file = temp, col_names = c("field", "year", "id_name"))

unlink(temp)
```

## Challenges

1. Take Chemistry (`Che`). Who received the most nominations?

2. Find all researchers who received nominations in more than one
   field.

3. Take Physics (`Phy`). Which year had the largest number of nominees?

4. For each field, what is the average number of nominees per year?
   Calculate the number of nominees in each field and year, and take
   the average across all years.

```{r Challenge part 1}
#Who has received the most nominations?
#Strategy: Filter field to only contain `Che`, group by author and count

nobel_nominations %>%
  filter(field == "Che") %>%
  group_by(id_name) %>%
  summarise(total_nominations = n()) %>% 
  ungroup() %>%
  arrange(desc(total_nominations)) %>%
  head(20)

```

```{r Challenge part 2}
##Which researchers have nominations in more than one field?

nobel_nominations %>%
  select(field, id_name) %>%
  distinct() %>%
  group_by(id_name) %>%
  mutate("total_fields" = n()) %>%
  ungroup() %>%
  filter(total_fields > 1) %>% 
  select(id_name, total_fields) %>%
  distinct() %>% view() %>%
  nrow()

#Answer is 177
```

```{r Challenge part 3}
#Which year had the largest number of Physics nominees?

nobel_nominations %>%
  select(field, year) %>% 
  filter(field == "Phy") %>%
  group_by(year) %>%
  mutate(nom_year = n()) %>%
  ungroup() %>%
  distinct() %>%
  arrange(desc(nom_year)) %>%
  head(20)

#Answer is 1957 and 1963
```

```{r Challenge part 4}
#For each field, what is the average number of nominees per year?

nobel_nominations %>%
  select(field, year) %>%
  group_by(field, year) %>%
  summarise("nom_field_year" = n()) %>%
  ungroup() %>%
  group_by(field) %>% 
  mutate("avg_nom_field_year" = mean(nom_field_year)) %>%
  select(field, avg_nom_field_year) %>%
  distinct() %>%
  head(5)

```

